---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. 

```{r}

# Your answer here.

```

2E2.

```{r}

# Your answer here.

```

2E3.

```{r}

# Your answer here.

```

2E4.

```{r}

# Discuss in pairs/groups. Ask me for a hint if you feel like it.

```

### Medium

2M1.

2M1. Recall the globe tossing model from the chapter. Compute and plot the grid approximate posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p.
(1) W,W,W
(2) W,W,W,L
(3) L,W,W,L,W,W,W

```{r}

# Write your code here.
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(4, size = 7, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")


```

2M2.
Now assume a prior for p that is equal to zero when p < 0.5 and is a positive constant when p ≥ 0.5. Again compute and plot the grid approximate posterior distribution for each of the sets of observations in the problem just above.

```{r}

# In the book, McElreath says that we can do much better than choosing a uniform prior for estimating the coverage of water. Is this what we are doing here? How does that impact our estimation?
prob_vec <- seq(from = 0, to = 1, length.out = 20)

prior_vec <- ifelse(p_grid < 0.5 , 0 , 0.6 )

likelihood <- dbinom(3, size = 4, prob = prob_vec)

unstd.posterior = prior_vec * likelihood

posterior = unstd.posterior / sum(posterior)

plot(prob_vec, posterior, type = "b")

```

2M3.

```{r}
# Use the Bayes formula!
# Write your probability here.
p = 0.3 * 0.5 / (0.3*0.5 + 1*0.5)

```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4.

```{r}




```

2M5.

```{r}

# Write your probability here.

```

2M6.

```{r}

# Write your probability here.

```
